**[Previous Section](multi_column_transformations.md)** | **[Index](../../README.md)** | **[Next Section](transformations_exercises.md)**
  
```{r, echo = FALSE, message = FALSE}
set.seed(1337)
library(mungebitsTransformations)
```

Dataframe Transformation
========
  
If the operation we are performing on our `data.frame` does not make column-by-column or row-by-row 
modifications, we are dealing with a whole dataframe transformation. For example, transposing our
`data.frame` using the built-in `t` function is such an operation.

In general, whole dataframe transformations are rare and should be broken down into compositions
of column and row transformations whenever possible. For example, "take the square root of each
entry" is a whole dataframe transformation, but it is *also* a column transformations:

```{r}
column_transformation(sqrt)(iris, 1:4)
t(head(cbind(datasets::iris[[1]], iris[[1]])))
```

Whenever possible, we should break down the operations we are performing on our data set
into simpler atomic chunks so that these chunks can be re-used and tested separately. If you
find yourself using whole data transformations, ask yourself if it can be achieved using a
composition of row and column transformations.

Exercises
--------

   1. Can re-ordering a `data.frame`'s columns (e.g., sorting columns by names) be accomplished
      using just row and column transformations? Why or why not?
      
   2. The following code removes all rows and columns containing more than 20% missing values.
      How would you break it down into row and column transformations?
      
```{r}
# Generate some NAs in iris
iris[] <- lapply(iris, function(x) ifelse(runif(150, 0, 1) > 0.7, NA, x))
remove_missings <- function(dataframe) eval.parent(substitute({
  dataframe <- dataframe[apply(dataframe, 1, function(x) mean(is.na(x)) <= 0.2), ]
  dataframe <- dataframe[, apply(dataframe, 2, function(x) mean(is.na(x)) <= 0.2)]
}))
remove_missings(iris)
dim(iris) # 76 4
```

**[Previous Section](multi_column_transformations.md)** | **[Index](../../README.md)** | **[Next Section](transformations_exercises.md)**